---
title: "Lab_10"
author: "CP"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos =c(CRAN = "http://cran.rstudio.com"))
```

```{r}
if (!require(RSQLite)) install.packages(c("RSQLite"))
if (!require(DBI))     install.packages(c("DBI"))                                

library(RSQLite)
library(DBI)


# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")
```

```{r}
# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```
#```{r}

#PRAGMA table_info(actor)
#```

```{r}
dbGetQuery(con, "PRAGMA table_info(actor)")
```
Exercise 1
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.



```{r}
dbGetQuery(con, "
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name")

```

Exercise 2
Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{r}
dbGetQuery(con, "
SELECT actor_id, first_name, last_name
FROM  actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
ORDER BY last_name
")

```

Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.
```{r}
dbGetQuery(con, "
SELECT DISTINCT customer_id , rental_date
FROM rental
WHERE date(rental_date) = '2005-07-05'")

```

```{r}
dbGetQuery(con, "
SELECT customer_id,
COUNT(*) AS N 
FROM rental
WHERE date(rental_date) = '2005-07-05'
GROUP BY customer_id")

```
Exercise 4
Exercise 4.1
Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{r}
dbGetQuery(con, "PRAGMA table_info(payment)")
```



```{r}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)")

```
Exercise 4.2
Construct a query that retrives all rows from the payment table where the amount is greater then 5

```{r}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount > 5 ")
```
Exercise 4.2
Construct a query that retrives all rows from the payment table where the amount is greater then 5 and less then 8

```{r}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount > 5 and amount < 8 ")
```
Exercise 5
Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.

```{r}
dbGetQuery(con, "PRAGMA table_info(customer)")
```

```{r}
dbGetQuery(con, "
SELECT c.customer_id, c.last_name, p.payment_id, p.amount 
FROM customer AS c INNER JOIN payment AS p 
ON c.customer_id = p.customer_id 
WHERE c.last_name == 'DAVIS'
 ")
```
Exercise 6
Exercise 6.1
Use COUNT(*) to count the number of rows in rental
```{r}
dbGetQuery(con, "
SELECT COUNT (*) AS count
FROM rental
")

```

Exercise 6.2
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT (*) AS count
FROM rental
GROUP BY customer_id 
")

```
Exercise 6.3 

```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT (*) AS count
FROM rental
GROUP BY customer_id 
ORDER BY count DESC

")

```
Exercise 6.4
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT (*) AS count
FROM rental
GROUP BY customer_id 
HAVING count >40 
ORDER BY count DESC
LIMIT 8

")
```


```{r}
dbGetQuery(con, "
SELECT MAX(amount) AS maxpayment,
       MIN(amount) AS minpayment,
       AVG(amount) AS avgpayment,
       SUM(amount) AS sumpayment FROM payment 

")
```
```{r}
dbGetQuery(con, "
SELECT customer_id, 
MAX(amount) AS maxpayment,
       MIN(amount) AS minpayment,
       AVG(amount) AS avgpayment,
       SUM(amount) AS sumpayment 
       FROM payment 
       GROUP BY customer_id
       LIMIT 5

")
```
```{r}
dbGetQuery(con, "
SELECT customer_id, 
COUNT(*) AS N, 
MAX(amount) AS maxpayment,
       MIN(amount) AS minpayment,
       AVG(amount) AS avgpayment,
       SUM(amount) AS sumpayment 
       FROM payment 
      
       GROUP BY customer_id
       HAVING N > 5
       LIMIT 5

")
```

```{r}
# clean up
dbDisconnect(con)
```


