---
title: "Lab 05"
author: "CP"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Read in the data
First download and then read in with data.table:fread()

```{r read-data, cache=TRUE}
if (!file.exists("../lab03/met_all.gz"))
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)


```
```{r}
met <- data.table::fread("met_all.gz")
```
```{r library}
library(tidyverse)
library(lubridate)
library(data.table)
library(dtplyr)
```


Read in the stations data 

```{r}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```


Merge met data with stations 

```{r merge datatables}
merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  ) 

```


```{r}
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

Remove temperatures less than -17C and change elev 9999 to missing value 
```{r remove-low-temp}
met <- met[temp > -17][elev ==9999.0, elev:=NA]
```
Question 1. 

Compute mean temperature, wind speed and 
```{r}
station_averages <- met[,. (temp= mean(temp, na.rm=T),
                            wind.sp = mean(wind.sp, na.rm=T),
                            atm.press= mean(atm.press, na.rm=T)), by = USAFID]
```

The above computes the mean by weather station. 

##Question 2 Now lets compute the median value for each variable 
```{r}
stmeds <- station_averages[,. (
                            temp50     = median(temp, na.rm=T),
                            windsp50   = median(wind.sp, na.rm=T),
                            atmpress50 = median(atm.press, na.rm=T)
)]
```

```{r}
station_averages[, temp_dist50 := abs(temp - stmeds$temp50)][order(temp_dist50)]

head(station_averages)
```

lets use 'which.min()'
It summarizes the min from above 

## Question 2

```{r}
merge(x= station_averages,
      y= stmeds,
      by.x = "STATE",
      all.x=T,
      all.y = FALSE)
```


```{r}
station_averages[,. (
                            temp_dist50     := (temp, na.rm=T),
                            windsp50   = median(wind.sp, na.rm=T),
                            atmpress50 = median(atm.press, na.rm=T)
```





```{r}
summary(met$elev)
```
Generate a date variable 

```{r create-ymd,cache=TRUE}
met <- met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
```
Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <-met[week(ymd)==31]
```
Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

```{r}
met[,.(
  temp     = max(temp,na.rm=TRUE),
  rh       = max(rh,na.rm=TRUE),
  wind.sp  = max(wind.sp,na.rm=TRUE),
  vis.dist = max(vis.dist,na.rm=TRUE),
  lat      = max(lat ,na.rm=TRUE),
  lon      = max(lon ,na.rm=TRUE), 
  elev     = max(elev,na.rm=TRUE)
)]
```
Great! No more 9999s in our dataset. 


```{r}
met_avg <- met[,.(
  temp     = mean(temp,na.rm=TRUE),
  rh       = mean(rh,na.rm=TRUE),
  wind.sp  = mean(wind.sp,na.rm=TRUE),
  vis.dist = mean(vis.dist,na.rm=TRUE),
  dew.point= mean(dew.point, na.rm=TRUE),
  lat      = mean(lat),
  lon      = mean(lon), 
  elev     = mean(elev,na.rm=TRUE)
), by=c("USAFID")]
```
Create a region variable for NW, SW, NE, SE based on lon = -93. 8.00 and lat = 39.71 degrees
```{r}
met_avg[, region := 
                fifelse(lon >=-98 & lat > 39.71, "NE",
                fifelse(lon <-98 & lat > 39.71, "NW",
                fifelse(lon <-98 & lat <= 39.71, "SW","SE")))]

 table(met_avg$region)                     
```

Create a categorical variable for elevation as in the lecture slides

```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]

```
## 3. MAke Violin plots of dew point temp by region 
