---
title: "Assignment_4"
author: "CP"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

### SQL

```{r}
library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```
```{r}
dbGetQuery(con, "PRAGMA table_info(category)")
```


```{sql, connection=con, output.var = "x1"}
PRAGMA table_info(film)

```

## Question 1
How many many movies is there avaliable in each rating catagory.

```{r}
dbGetQuery(con,"
SELECT  rating, COUNT(*) AS count
FROM  film
GROUP BY rating
")
```

## Question 2
What is the average replacement cost and rental rate for each rating category.
```{r}
dbGetQuery(con,"
SELECT  rating, replacement_cost, rental_rate,
COUNT(*) AS count,
AVG(replacement_cost) AS avgreplacement_cost
FROM  film
GROUP BY rating
")
```

## Question 3
Use table film_category together with film to find the how many films there are with each category ID

```{r}
dbGetQuery(con,"
SELECT  film.film_id, film_category.category_id,
COUNT(*) AS N
FROM film
INNER JOIN film_category  ON film.film_id = film_category.film_id
GROUP BY category_id
")


```


## Question 4 
Incorporate table category into the answer to the previous question to find the name of the most popular category.

```{r}
dbGetQuery(con,"
SELECT  film.film_id, film_category.category_id,category.category_id, category.name,
COUNT(*) AS N
FROM film
INNER JOIN film_category  ON film.film_id = film_category.film_id
INNER JOIN category ON category.category_id = film_category.category_id
GROUP BY name 
")
```
Category 15 which is Sports is the most popular category with 74 movies. 

### HPC

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}
```

```{r}
fun1alt <- function(mat) {
  rowSums(mat)
}

```

```{r}
# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)
```

```{r}
# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat))
```


```{r}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}
```

```{r}
fun2alt <- function(mat) {
  cumsum(mat) }


```

```{r}
# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat))
```


```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132

# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```



```{r}
library(parallel)
```

```{r}
system.time({
  cl <- parallel::makePSOCKcluster(2L)

parallel::clusterSetRNGStream(cl, 123)
  ans <- unlist(parallel::parLapply(cl, 1:4000, sim_pi, n = 10000))
  print(mean(ans))
  parallel::stopCluster(cl)
})

```





