---
title: "Lab 03"
author: "CP"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Read in the data
First download and then read in with data.table:fread()

```{r read-data, cache=TRUE}
if (!file.exists("../lab03/met_all.gz"))
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)


```
```{r}
met <- data.table::fread("../lab4/met_all.gz")
```
```{r library}
library(tidyverse)
library(lubridate)
library(data.table)

```

## Step 2 Prep data 

Remove temperatures less than -17C and change elev 9999 to missing value 
```{r remove-low-temp}
met <- met[temp > -17][elev ==9999.0, elev:=NA]
```
```{r}
summary(met$elev)
```
Generate a date variable 

```{r create-ymd,cache=TRUE}
met <- met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
```
Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <-met[week(ymd)==31]
```
Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

```{r}
met[,.(
  temp     = max(temp,na.rm=TRUE),
  rh       = max(rh,na.rm=TRUE),
  wind.sp  = max(wind.sp,na.rm=TRUE),
  vis.dist = max(vis.dist,na.rm=TRUE),
  lat      = max(lat ,na.rm=TRUE),
  lon      = max(lon ,na.rm=TRUE), 
  elev     = max(elev,na.rm=TRUE)
)]
```
Great! No more 9999s in our dataset. 


```{r}
met_avg <- met[,.(
  temp     = mean(temp,na.rm=TRUE),
  rh       = mean(rh,na.rm=TRUE),
  wind.sp  = mean(wind.sp,na.rm=TRUE),
  vis.dist = mean(vis.dist,na.rm=TRUE),
  dew.point= mean(dew.point, na.rm=TRUE),
  lat      = mean(lat),
  lon      = mean(lon), 
  elev     = mean(elev,na.rm=TRUE)
), by=c("USAFID")]
```
Create a region variable for NW, SW, NE, SE based on lon = -93. 8.00 and lat = 39.71 degrees
```{r}
met_avg[, region := 
                fifelse(lon >=-98 & lat > 39.71, "NE",
                fifelse(lon <-98 & lat > 39.71, "NW",
                fifelse(lon <-98 & lat <= 39.71, "SW","SE")))]

 table(met_avg$region)                     
```

Create a categorical variable for elevation as in the lecture slides

```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]

```
## 3. MAke Violin plots of dew point temp by region 

```{r violin-temp}
met_avg[!is.na(region)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(y =temp, x = 1, color=region, fill=region)) + 
  facet_wrap(~ region, nrow = 1)
```
```{r}
met_avg[, region := 
                fifelse(lon >=-98 & lat > 39.71, "NE",
                fifelse(lon <-98 & lat > 39.71, "NW",
                fifelse(lon <-98 & lat <= 39.71, "SW","SE")))]

 table(met_avg$region)
```
```{r violin dew.point}

met_avg[!is.na(region)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(y =dew.point, x = 1, color=region, fill=region)) + 
  facet_wrap(~ region, nrow = 1)
```
the highest dew point temperatures recorded in the south west. 
```{r violin wind.sp}

met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(y =wind.sp, x = 1, color=region, fill=region)) + 
  facet_wrap(~ region, nrow = 1)
```
the highest wind speed occured in the Northeast 

##4. Use geom_point and geom_smooth to examine the association between dew point temperature and wind speed by region

```{r scatterplot-dewpoint-winds.p}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x =wind.sp, y = dew.point, color=region,)) + 
  geom_point() + 
  geom_smooth(method = lm, mapping = aes(linetype = region, color ="black"))+
  facet_wrap(~ region, nrow = 1)

```
Southwest is the only plot decreasing. 

##5. Use geom_bar to creat bar plots of the weather stations by elevation category 
```{r}

met_avg %>%
filter(!(region %in% NA)) %>% 
  ggplot(mapping = aes(x = elev)) + 
  geom_bar(position="dodge")
```
```
## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot (mapping = aes(x = region, y = dew.point))
    stat_summary(fun.data= mean_sdl)
```



