---
title: "Lab 03"
author: "CP"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Read in the data
First download and then read in with data.table:fread()

```{r}
if (!file.exists("../lab03/met_all.gz"))
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)


```
```{r}
met <- data.table::fread("../lab4/met_all.gz")
```
```{r library}
library(tidyverse)
library(lubridate)
library(data.table)

```

## Step 2 Prep data 

Remove temperatures less than -17C and change elev 9999 to missing value 
```{r remove-low-temp}
met <- met[temp > -17][elev ==9999.0, elev:=NA]
```
```{r}
summary(met$elev)
```
Generate a date variable 

```{r create-ymd}
met <- met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
```
Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <-met[week(ymd)==31]
```
Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

```{r}
met[,.(
  temp     = max(temp,na.rm=TRUE),
  rh       = max(rh,na.rm=TRUE),
  wind.sp  = max(wind.sp,na.rm=TRUE),
  vis.dist = max(vis.dist,na.rm=TRUE),
  lat      = max(lat ,na.rm=TRUE),
  lon      = max(lon ,na.rm=TRUE), 
  elev     = max(elev,na.rm=TRUE)
)]
```
Great! No more 9999s in our dataset. 


```{r}
met_avg <- met[,.(
  temp     = mean(temp,na.rm=TRUE),
  rh       = mean(rh,na.rm=TRUE),
  wind.sp  = mean(wind.sp,na.rm=TRUE),
  vis.dist = mean(vis.dist,na.rm=TRUE),
  lat      = mean(lat),
  lon      = mean(lon), 
  elev     = mean(elev,na.rm=TRUE)
), by=c("USAFID")]
```
















Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
Create a categorical variable for elevation as in the lecture slides



