---
title: "Assignment_02"
author: "Camille Parchment"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
if (!file.exists("individual.csv"))
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", "individual.csv", method="libcurl", timeout = 60)
```
```{r}
indv <- read.csv("individual.csv")
```

```{r}
if (!file.exists("regional.csv"))
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", "regional.csv", method="libcurl", timeout = 60)
```
```{r}
reg <- read.csv("regional.csv")
```

### Merge data 

```{r}
chs <- merge(indv, reg, by="townname")
```
```{r}
nrow(chs)
```

After merging based on  location (townname) there are still 1200 observations in our data set.

```{r}
library(tidytext)
library(tidyverse)
library(data.table)
library(dplyr)
library(ggplot2)
```
```{r}
chs <- as_tibble(chs)
```
```{r}
library(dtplyr)
```
### Impute data using the average within the variables “male” and “hispanic.”

```{r}
chs[, bmi_imp = fcoalesce(bmi, mean(bmi, na.rm = TRUE)),
by = .(male,hispanic)]

```

### Create a BMI catergorical variable and a summary table with underweight, normal weight, overweight, and obese. 

```{r}
obesity_level <- cut(chs$bmi, breaks=c(0, 14, 22, 24, 100), labels= c("underweight","normal", "overwight","obese"), right = FALSE)
```
```{r}
obesity_level[1:10]
```
```{r}
chs$bmi[1:10]
```
```{r}
summary(obesity_level) 
```
```{r}
summary(chs$bmi, na.rm = TRUE)
```



```{r}
smoke_gas_exposure <- ifelse(chs$smoke == 0 & chs$gasstove == 0 , "none",
                                            ifelse(chs$smoke == 1 & chs$gasstove == 0, "smoke_only",
                                            ifelse(chs$smoke == 0 & chs$gasstove == 1, "gas_only",
                                            ifelse(chs$smoke == 1 & chs$gasstove == 1, "both", NA))))
                                            table(smoke_gas_exposure, useNA = "always")
```
```{r}
library(data.table)
```
```{r}
chs <- data.table(chs)
```
### 4 tables gouped by town, sex, obesity_level, and smoke gas exposure
```{r}
town_averages <- chs[, .(
    fev_avg      = mean(fev, na.rm=TRUE),
    asthma_avg   = mean(asthma, na.rm=TRUE)),
    by = townname
    ][order(townname)] %>% head(n=10)
```
```{r}
sex_averages <- chs[, .(
    fev_avg      = mean(fev, na.rm=TRUE),
    asthma_avg   = mean(asthma, na.rm=TRUE)),
    by = male
    ][order(male)] %>% head(n=10)
```

```{r}
Obs_averages <- chs[, .(
    fev_avg      = mean(fev, na.rm=TRUE),
    asthma_avg   = mean(asthma, na.rm=TRUE)),
    by = obesity_level
    ][order(obesity_level)] 
```

```{r}
head(town_averages)
```
```{r}
head(sex_averages)
```
```{r}
head(Obs_averages)
```
### Part 2
### 1. facet plot of BMI and FEV by townname 
```{r}
ggplot(chs, aes(x= bmi, y= fev, color= townname)) + geom_point() + facet_wrap(~ townname, ncol = 4) + labs(title = "Bmi by Forced Expiratory Volume Grouped by Town Name", x= "BMI", y="forced expiratory volume")
```
### 2. Stacked Histogram of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r}
ggplot(chs, aes(x=fev, fill= obesity_level)) + geom_histogram() + scale_fill_brewer()
ggplot(chs, aes(x=fev, fill= smoke_gas_exposure)) + geom_histogram() + scale_fill_brewer()
```

### 3. Barchart of BMI by smoke/gas exposure.

```{r}
ggplot(chs, aes(x= obesity_level, fill= smoke_gas_exposure )) +geom_bar() + labs(title= "Obesity level by Smoke, Gas or Both", x="Obesity category", y="Count")
```
### 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category
 
```{r}
ggplot(chs, aes(x= obesity_level, y= fev, fill= obesity_level))+ geom_boxplot() + labs(title= "Forced Expiratory Volume by BMI", x= "obesity catergory", y= "forced expiratory level")
ggplot(chs, aes(x=smoke_gas_exposure, y= fev, fill= smoke_gas_exposure)) + geom_boxplot() + labs(title = "Forced Expiratory Volume by Smoke/Gas Exposure", x= "smoke gas exposure", y= "forces expiratory volume")
```

### A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
 
```{r}
library(leaflet)
```

```{r}
pm2.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=chs$pm25_mass)

```

```{r}
Pm25mass_map <- leaflet(chs) %>%
      addProviderTiles('CartoDB.Positron') %>% 
      addCircles(
    lat = ~lat, lng=~lon,
                                                  
    label = ~paste0(townname), color = ~ pm2.pal(pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>% addLegend('bottomleft', pal= pm2.pal, values=chs$pm25_mass,
          title='Concentrations.', opacity=1)
```


```{r}
pm25_avg <- chs[, .(
    pm25_mass    = mean(pm25_mass, na.rm=TRUE)),
    by = townname
    ][order(townname)] 
```

### 6. Scatter plot of Pm25 mass by Forced expiratory volume by town name 

```{r}
ggplot(chs, aes(x= pm25_mass, y= fev, color= townname)) + geom_point() + labs(title= "Pm 2.5 mass vs Forced expiratory volume", x= "Pm 2.5 mass", y= "Forced Expiratory Volume")

```



```{r}
RColorBrewer::display.brewer.all()
```







